Manuscript_draft

### 2021-01-08 manuscript_draft.Rmd
manny work

## Tree-level growth responses

@@

## Stand-level growth and structure

### Stand-level volume and basal area estimates
Total and merchantable volumes were estimated for trees using species- and region-specific equations [@nigh2016] based on tree height and DBH.  Consistent with forest management practices in British Columbia, we estimated volumes for trees meeting a minimum diameter utilization limit of 18.5cm DBH.  For trees lacking a tree height measurement, we estimated height by fitting a single diameter-height model developed with a linear mixed effects model, with log-transformed DBH as a fixed effect, and tree nested within plot as a random effect.  Trees with broken tops were omitted from diameter-height models. We summed tree-level estimates of volume and basal area to generate stand-level estimates for each plot. To test the hypothesis that changes in stand attributes over 27 years varied by treatment, we fit separate linear models with stand total volume, basal area, and density as response variables, and the interaction between treatment and year as the independent variable.

### Mortality

@@

The tree-level basal area increment model indicated that interactions between log-transformed tree size, species and treatment explained `r round(r.squaredGLMM(fitBA)[2]*100,1)`% of the variation in log-transformed 1992-2019 tree-level BAI (`r tableNums("fitBA.table",display="cite")`).  Estimated marginal means from the model showed that between species and among treatments, larger trees increased basal area more than smaller trees (`r figNums("treeGrowth.plot",display="cite")`).  Species contrasts showed that spruce trees increased  basal area more than than fir in the high-removal treatment unit, whereas basal growth between the two species was close to equal in the low-removal and control units (`r appNums("sppContrast.table",display="cite")`).  Within each species, treatment contrasts showed that spruce basal area increment in the high-removal unit exceeded spruce growth in both other units (`r appNums("trtContrast.table",display="cite")`), whereas fir basal area increment in the high- and low-removal units was higher than the control unit, and there were no significant differences between the two harvested units (`r appNums("trtContrast.table",display="cite")`).

@@

appNums("htModelFit.plot",display="FALSE") # model output

# Tree heights model
  appNums("heightDiam.plot",display=FALSE) # Plot of diameter/heights`

@@

`r figNums("treeBAI.plot",caption="Estimated marginal means of spruce and fir basal area increment (m²) across five inital DBH classes and three treatments.  Whiskers are standard error of the mean.  Responses are back-transformed from the model.")`

@@

```{r function to create figure of plot means, fig.height=10}

@@

>>>>

## Data analyses

### Tree-level growth responses

@@

### Stand-level growth and structure

#### Stand-level volume and basal area estimates
Total and merchantable volumes were estimated for trees using species- and region-specific equations [@nigh2016] based on tree height and DBH.  Consistent with forest management practices in British Columbia, we estimated volumes for trees meeting a minimum diameter utilization limit of 18.5cm DBH.  For trees lacking a tree height measurement, we estimated height by fitting a single diameter-height model developed with a linear mixed effects model, with log-transformed DBH as a fixed effect, and tree nested within plot as a random effect.  Trees with broken tops were omitted from diameter-height models. We summed tree-level estimates of volume and basal area to generate stand-level estimates for each plot. To test the hypothesis that changes in stand attributes over 27 years varied by treatment, we fit separate linear models with stand total volume, basal area, and density as response variables, and the interaction between treatment and year as the independent variable.

#### Mortality

@@

The tree-level basal area increment model indicated that interactions between log-transformed tree size, species and treatment explained `r round(r.squaredGLMM(fitBA)[2]*100,1)`% of the variation in log-transformed 1992-2019 tree-level BAI (`r tableNums("fitBA.table",display="cite")`), and `r round(r.squaredGLMM(fitBApct)[2]*100,1)`% of the variation in basal area increment as a percentage of tree size in 1992 (`r tableNums("fitBApct.table",display="cite")`).  Estimated marginal means from the model showed that between species and among treatments, basal area increment increased with tree size, meaning that larger trees increased basal area more than smaller trees (`r figNums("treeGrowth.plot",display="cite")`), although smaller trees increased their size proportionally more than large trees (`r figNums("treeGrowthPct.plot",display="cite")`).  Species contrasts showed that spruce trees increased  basal area more than than fir in the high-removal treatment unit, whereas basal growth between the two species was close to equal in the low-removal and control units (`r appNums("sppContrast.table",display="cite")`).  Within each species, treatment contrasts showed that spruce basal area increment in the high-removal unit exceeded spruce growth in both other units (`r appNums("trtContrast.table",display="cite")`), whereas fir basal area increment in the high- and low-removal units was higher than the control unit, and there were no significant differences between the two harvested units (`r appNums("trtContrast.table",display="cite")`).

@@

@@

`r figNums("treeGrowth.plot",caption="Estimated marginal means of spruce and fir basal area increment (m²) across five inital DBH classes and three treatments.  Whiskers are standard error of the mean.  Responses are back-transformed from the model.")`

@@

```{r tree BAI as percentage by treatment species and diameter class}

# Using estimated marginal means
  emmeans(rg.pbaipct,~Init|BA.Target+Species,type="response") %>% 
  as.data.frame() %>%
  mutate(Init=ba2d(Init)) %>% # convert initial basal area back to diameter

  # plotting
  ggplot()+
  aes(x=Init,y=response,fill=Species,color=Species)+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(0.05))+
  xlab("1992 DBH (cm)")+
  ylab("Basal area increment from 1992-2019 (% increase)")+
  facet_wrap(~BA.Target)+
  theme(legend.position="bottom")

`r figNums("treeGrowthPct.plot",caption="Estimated marginal means of spruce and fir basal area increment (expressed as a percentage increase relative to 1992 DBH) across five inital DBH classes and three treatments.  Whiskers are standard error of the mean.  Responses are back-transformed from the model.")`

```{r plot means plots, fig.height=10}

@@

```{r plot means by species}

plotSums<-
  plotTotal.X %>% 
  group_by(BA.Target,Plot,Year,Var) %>% 
  summarise(Value=sum(Value)) %>% 
  mutate(Species="Total") %>% 
  dplyr::select(BA.Target,Plot,Year,Var,Species,Value)

plotTotal.X %>%

  # bring in plotSums
  dplyr::select(BA.Target,Plot,Year,Var,Species,Value) %>% 
  bind_rows(plotSums) %>%

  # need to convert year to numeric for plotting
  mutate(Year=as.numeric(as.character(Year))) %>%

  group_by(BA.Target,Year,Species,Var) %>% 
  summarise(Mean=mean(Value),
            SE=sd(Value)/sqrt(length(Value))) %>%

  ggplot()+
  aes(x=Year,y=Mean,group=Species,fill=Species,color=Species)+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE, width=.2),
                   position=position_dodge(0))+
 scale_x_continuous(limits=c(1992,2019),breaks=c(1992,1994,1997,2009,2019),
                     minor_breaks = c(1992,1994,1997,2009,2019),
                     name="Measurement year")+ # stretch out x-axis

  facet_grid(Var~BA.Target,
             scales="free_y")+
  theme(legend.position = "bottom",
  axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5))
